Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey

### 2014-01-21 structured.tex
2012.06.11 (Vassilis) Content revisited. Duplicate text has been removed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Geographic Layout}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Mithos}
\textit{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which
incorporates directed incremental probing to find near optimal node placement,
and is classified by its authors as an integration of geographic layout and
proximity routing overlay optimization methods.

The bootstrap phase of \textit{Mithos} starts with a subset of existing members
as the first set of candidates, and while the iteratively closest nodes are
detected by probing the neighborhood, the candidate neighbor list is updated. In
order to avoid a local minima \textit{Mithos} probes all the neighbours within a
two hop distance from the current minimum before concluding the process.

@@

The last step of the algorithm is the interconnection among neighbours.
\emph{Mithos} uses a \emph{quadrant}-based mechanism according to which each
node establishes a link to the closest neighbour in each quadrant. During
forwarding, the next hop is performed towards a neighbour in the same quadrant
as the final destination. The newcoming node may not know of other neighbours
in all quadrants, therefore, the node first identifies neighbours in all
quadrants using a mechanism based on ideas similar to a perimeter
walk\footnote{Used in Greedy Perimeter Stateless Routing (GPSR) protocol.} and
then improves the results using parallel path processing by taking into account
further geometric properties of node relationships.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf LAPTOP}
\emph{LAPTOP} \cite{wu_laptop_2007},  organises the overlay into a tree-based
hierarchy with main focus on reducing hops during message routing as well as
minimizing maintenance overhead. Additionally, a caching scheme is also
incorporated so as to further reduce routing table update costs. The authors
theoretically show that in \textit{LAPTOP} routing path length is bounded by
$O(log_d N)$ and node joining and leaving in the overlay network is bounded by
$O\left( d log_d N \right)$ hops in a balanced overlay tree, where $N$ is the
number of nodes, and $d$ is the maximum degree of each node. \textit{LAPTOP}
implements the geographical layout approach  and constructs a geographical
layout in a self-organizing and efficient fashion, by estimating the
round-trip-times to a small number of nodes in the overlay network in order to
make them roughly aware of their physical distances among them.

Each node is assigned an address in a dotted format (e.g 1.3.4). Each octet
ranges from $1$ to $d$, where $d$ is the maximum degree of the nodes. The
assignment process is done by appending a unique octet to the address of each
nodes
parent, while root node is assigned address 1.  The routing scheme is similar to
the longest-prefix IP matching scheme. At each forwarding hop, any message
travels up the tree until the first common ancestor of both source and
destination nodes is reached and then starts descending to arrive to its target.
During tree traversals, special entries in the routing tables, called
\emph{routing cache}, are maintained in order to increase routing efficiency and
achieve finer load balance. Caching enables a node to forward a message to a
better longest-prefix match than that of its direct ancestor making a larger,
quicker and more cost effective step through the overlay and toward the
destination. To improve scalability, the number of children nodes and the size
of the routing cache are limited.  In terms of overlay maintenance, LAPTOP
incorporates a simple \emph{heartbeat}-based technique where each parent node is
responsible for monitoring its children.  At join process the newcomming node is
assigned its level label as well as its address by its parent node. Additionally
it initializes its routing table (with normal and caching entries) as it
traverses the overlay in search for its parent node.  During a graceful
departure, the referred node checks for children in the overlay. If it does not
have any, it simply notifies its parent and leaves. If it has, it selects the
child node with the lowest round-trip-time to it in order to take its place so
that the locality property is preserved.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Proximity Routing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Proximity routing does not require routing tables to be built using any
knowledge about network proximity. On the other hand it exploits such knowledge
in order to choose the best next hop during routing a message. This approach
balances between choosing the node that will further progress the routing
towards the destination and choosing the closest entry in the routing table, in
terms of network proximity.
% TODO: looks like discussion material
Thus, it is relatively less effective than
geographical layout when applied to CAN(-like) implementations. Moreover, the
technique has been incorporated into a version of Chord causing an increase on
the overhead of node joins and the size as well as maintenance cost of finger
tables.
%
%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
%the domain of proximity routing selection. In \emph{CFS}, each node predicts
%the entire lookup latency as a function of the total number of nodes and the
%average overlay next routing peer. The problem is that it is very difficult to
%have a clear picture on the total number of nodes and the average hop latency
%from the local. This leads to rough estimations that consequently decreases
%overall performance.
%
The methods that use proximity routing and their details are listed below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Proximity in Kademlia}
Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
summarize the highlights about the {\em Kademlia} protocol. {\em Kademlia}
\cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
protocol for P2P networks, which uses an iteration based lookup algorithm. The
protocol uses the standard $160$ bit ID system for nodes and locates the nodes
in a prefix binary tree, where IDs are used as prefixes. The iterative lookup
operations are done over this prefix binary tree, which converges to logarithmic
lookup times. The ID's are assigned randomly, therefore in Kademlia, there is no
proximity control, which results in inefficient use of the underlying network
during lookup and retrieve operations.

{\em Proximity in Kademlia} \cite{kaune_pkad_2008} introduces proximity controls
over the base Kademlia protocol to optimize the underlying network usage of the
protocol. The authors define an abstract {\em underlay metric} that calculates
the suitability of establishing a communication link between peers as a cost
function, based on the used proximity criteria (RTT, ISP locality, etc.).
{\em Proximity in Kademlia} adapted both {\em proximity routing} and {\em
proximity neighbour selection} overlay optimization algorithms for controlling
proximity. Therefore, we categorize the protocol only in this section, and skip
the discussion in the further {\em proximity neighbour selection} section
(Sec.\ref{sec:pns}). {\em Proximity in Kademlia} also uses the MaxMind GeoIP
database\footnote{MaxMind Geolocation Technology. http://www.maxmind.com} to
detect the proximity information of a given peer based on its region, country,
or ISP location, and use this information to form clusters of nearby peers. One
other method {\em Proximity in Kademlia} uses is the Vivaldi
\cite{cox_vivaldi_2004} protocol. However, authors report that clustering
approach performs better than Vivaldi protocol based on their experiments.
Authors also report that {\em Proximity routing} protocol successfully worked
with the Kademlia protocol and improved the locality of the connections over
peers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf CHOord considering Proximity on IPv6 (CHOP6)}
\textit{CHOP6} \cite{morimoto_chop6_2007} is designed based on the
\textit{Chord} protocol. \textit{CHOP6} roughly estimates the proximity among
nodes by exploiting the IPv6 address format and RTT information if available.
The proximity estimation is achieved by introducing a 64-bit ID scheme in which
the least significant bit part is the IPv6 global routing prefix and thus
enabling a longest prefix match scheme. The protocol is designed based on the
observation that it is possible to estimate a node's geographical location by
simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
\textit{Chord}, \textit{CHOP6} uses a finger table, whose entries hold more than
one candidate node.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Chord6}
\emph{Chord6} \cite{xiong_chord6_2005} is another \textit{Chord} variant that
tries to exploit the hierarchical features of IPv6 in order to create a
substrate that reduces interdomain traffic between service providers.
\textit{Chord6} is based on the original Chord protocol, and the main difference
from the original protocol is in the identifier definition. Therefore, the
approach can be easily portable to other DHTs such as CAN, Pastry and Tapestry.
In Chord6 the identifier contains two parts: the higher bits are obtained by
hashing the node's IPv6 address prefix of specific length, while the remaining
lower bits are the hash value of the rest of that IPv6 address. As a result of
this assignment, nodes in a domain will be mapped onto a continuous key space on
the overlay network, which avoids unnecessary message forwarding across
different service providers, thus minimizing overall routing cost.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Proximity Neighbour Selection}\label{sec:pns}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\textit{Proximity Neighbour Selection} constructs the routing tables using
proximity knowledge. The proximity information used in this method is different
than the landmark based systems described in the \textit{Geographic Layout}
section, as time-to-live values between nodes, or directly node ID prefixes are
used to detect proximity.  Tapestry, Pastry, and CAN successfully implemented
the proximity into their algorithms by using this approach.  The routing
protocol in Pastry is based on longest node ID prefix matching, while CAN uses
RTT values to detect close by nodes. \cite{castro_proximityp2p_2002}  reports
that proximity neighbour selection as an effective proximity based method.
Proximity Neighbor Selection based algorithms are described in detail below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf DHT-PNS}
Chord-DHT-PNS \cite{hancong_pnsbased_2006} implements proximity neighbour
selection on top of the Chord DHT. The main goal is to use proximity information
to group physically close by nodes as neighbours in the DHT table. In order to
detect proximity, virtual network coordinates of peers are used, by using the
Vivaldi protocol \cite{cox_vivaldi_2004}. The virtual coordinates are then used
by the nodes to map to identifier space in the DHT. The space is partitioned
using a \emph{concentric circle clustering scheme} where successive cycles of
radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. Then the formed
annuluses are divided into $2\chi-1$ \emph{sectors}, where $\chi$ denotes the
sequence number of the annulus starting from $\chi = 1$ for the centre cycle. It
is proved in the paper, that this way each sector occupies the same area as does
the center cycle. Assuming uniform node distribution, this characteristic,
favours a more load balanced clustering operation. Every sector in this $2D$
coordinate space is mapped to a unique \emph{region} in the DHT space forming a
multi-layer node identifier space.  Thus, any nodes that belong to the same
sector, are mapped to the same region as well, preserving their proximity
relationship unveiled by the use of the Vivaldi protocol. The individual pieces
are mapped to the identifier space uniquely, allowing logarithmic lookup
operations with high probability on Chord.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf IP-based Clustering (IPBC)}
Proximity neighbour selection algorithms use probing and other techniques
to detect proximity. However, such methods are either not precise enough, or
overload the network. \emph{IP-based clustering}
\cite{karwaczynski_ipbc_2007} is a proximity neighbour selection based
algorithm in which the authors propose to use the IP address prefixes (16 bit
for IPv4) in order to detect proximity.  \cite{freedman_iploc_2005} states that
$97\%$ of prefixes larger than $24$ bits belong to a single geographical
location. However, using less number of bits creates less precise results and
more number of bits increase the burden on the network and reduce the possible
number of neighbours. Therefore, a careful selection is required in terms of
performance/accuracy tradeoff. The {\em IP-based Clustering} generates a key and
its prefix and stores the later in the DHT itself, so that any newly joining
nodes with the same IP prefix can query it and identify all the
neighbours with the same one relatively easily. Nodes periodically update
their entries in the DHT by removing those who leave, or those who has been
timed out with no further activity detected from the node.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Cone}
\textit{Cone}\cite{wang_cone_2007} extends Chord using proximity neighbour
selection topology optimization algorithm. The proximity information is
generated using landmarks and round-trip-time-based distances to landmarks.
\textit{Cone} uses a two-layered identifier space. The first, named Chord-layer
identifier, denoted as $Id_{Chord}$, is the same as in Chord. The second is the
Cone-layer identifier, $Id_{Cone}$ which is constructed by two component
identifiers. The first, known as \emph{group identifier (gid)} denotes a
relevant group the node belongs to while the second, namely \emph{local
identifier (lid)} indicates the local identifier within the group. The group
concept, which is introduced here, is a way of dividing nodes according to a
common $Id_{Chord}$ prefix.  The structure of a Cone overlay, retains the
Chord's circular topology. The difference lies on the fact that, now, two rings
are created. A big ring, where nodes with the same \emph{gid} are arranged at
each position. Each of these positions are a smaller ring for the particular
group's \emph{lid}s. The routing is achieved in both clockwise and
counter-clockwise directions in the big-ring, for which two routing tables are
maintained, namely \emph{front} and \emph{back} finger tables. Entries in these
tables, display physical network proximity with the current node. Moreover, a
third table called \emph{group} maintains information about other online peers
within the current node's group in a way that entries are now close in the ID
space.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Geographic Layout} \label{section:geographic_layout}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% TODO: Geographic Layout
%
%\begin{itemize}
% \item Geographic Layout\\
%The node IDs are assigned in such a way that nodes close by in the physical
%network topology, be close in the node ID space as well. Implementations that
%work relatively well with this approach have been incorporated into CAN. Nodes
%measure the RTT between themselves and a set of landmarks in order to match the
%CAN space as much as possible to the physical one. Unfortunately, the approach
%requires well known landmark servers and for that matter is  not fully
%self-organizing which can further lead to imbalanced node distribution. On
%other DHTs, such as Chord or Pastry, another problem emerges.  To gain fault
%tolerant properties, these protocols, replicate key-value pairs on neighbouring
%(in the ID space) nodes. When a proximity-based node ID assignment has been
%used, the needed failure resilience is undermined by the fact that close by
%nodes are more likely to suffer collective failures.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%
% TODO: SOME DISCUSSION
%
% Xu \textit{et al.} \cite{xu_globstate_2003} state that there are three ways
% generating proximity information
%\begin{itemize}
% \item Expanding Ring Search\\
%Expanding ring search can be of two forms. First it can utilize the multicast
%infrastructure in the underlying network in order to emit its messages.
%Unfortunately such infrastructures are not widely deployed thus the
%implementations of this way of generating proximity information is limited to
%blindly flooding the neighbourhood to obtain reasonable results.
%
% \item Heuristics\\
%Heuristics are used in order to reduce the blindness of the expanding ring
%search and make realizations more efficient and effective. Unfortunately a
%common problem of all heuristic approaches is the local minimum pitfall in
%which
%the search might be caught into.
%
% \item Landmark Binning\\
%Landmark clustering is based on the view that nodes with similar distances to a
%set of predefined well-known landmark nodes are pretty likely of being close to
%each other. But this approach has its weaknesses as well, such as the fact that
%is a rather coarse grained approximation, therefore not particularly well
%suited
%for differentiating nodes within close distance to each other.
%
%\end{itemize}
%
%

%
% TODO: IN PAPER INTRODUCTION A DISCUSSION ABOUT
%TOPOLOGY-AWARE CAN
%
%Techniques to exploit topology information in overlay
%routing include geographic layout, proximity routing and
%proximity neighbor selection [3]. With geographic layout
%such as topology-aware CAN [12], the overlay structure is
%constrained by underlying network topology. This tech-
%nique, unfortunately, can create uneven distribution of
%nodes in the overlay, increasing the chances of overloading
%nodes and rendering the maintenance cost formidable. Our
%study shows that, for a typical 10,000-node topology-aware
%CAN, 5% nodes occupy 85-98% of the entire Cartesian
%space, and some nodes have to maintain 450-1500 neigh-
%bors. In Proximity routing, physical topology is not consid-
%ered when constructing the overlay.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Mithos}
\emph{Mithos} \cite{waldvogel_mythos_2003} is a P2P protocol which incorporates
directed incremental probing to find near optimal node placement, and is
classified by its authors as an integration of geographic layout and proximity
routing overlay optimization methods.

The bootstrap phase of Mithos starts with a subset of existing members as the
first set of candidates, and while the iteratively closest nodes are detected by
probing the neighborhood, the candidate neighbor list is updated. In order to
avoid a local minima, Mithos probes all the neighbours within a two-hop distance
from the current minimum before concluding the process.

@@

The last step of the algorithm is the interconnection among neighbours. Mithos
uses a \emph{quadrant-based} mechanism according to which each node establishes
a link to the closest neighbour in each quadrant. During forwarding, the next
hop is performed towards a neighbour in the same quadrant as the final
destination. The newcoming node may not know of other neighbours in all
quadrants, therefore, the node first identifies neighbours in all quadrants
using a mechanism based on ideas similar to a perimeter walk\footnote{Used in
Greedy Perimeter Stateless Routing (GPSR) protocol.} and then improves the
results using parallel path processing by taking into account further geometric
properties of node relationships.

@@

%
% TODO: review this part
%
%In order to avoid local minima during neighbour detection, extensive probing
%  must be undertaken. In simulation, unfortunately, only very small-sized
% overlay topologies (of 200 to 1000 nodes) have been used and thus no safe
% conclusions can be made as for the behaviour of an extensively large,
% real-world p2p deployment of the scheme.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf LAPTOP}
\emph{LAPTOP} \cite{wu_laptop_2007}, organises the overlay into a tree-based
hierarchy with main focus on reducing hops during message routing as well as
minimizing maintenance overhead. Additionally, a caching scheme is also
incorporated so as to further reduce routing table update costs. The authors
theoretically show that in LAPTOP routing path length is bounded by $O(log_d N)$
and node joining and leaving in the overlay network is bounded by
$O\left( d log_d N \right)$ hops in a balanced overlay tree, where $N$ is the
number of nodes, and $d$ is the maximum degree of each node. LAPTOP implements
the geographical layout approach  and constructs its layout in a self-organizing
and efficient fashion, by estimating the round-trip times to a small number of
nodes in the overlay network in order to make them roughly aware of their
physical distances among them.

Each node is assigned a dot-formated address (e.g 1.3.4). Each octet ranges from
$1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment process
is done by appending a unique octet to the address of each nodes parent, while
root node is assigned address $1$.

%The protocol is based on four definitions.
%\begin{enumerate}
% \item The amplitude of all possible measured RTTs is devided into intervals.
%  Each node measures its distance to its parent and is assigned a label $L_i$
% where $i$ denotes the configurable RTT interval in which the measured distance
% falls into. A special kind of node, the root, is initially assigned the $L_1$
% label and maintains (as all $L_1$ nodes do) a list of other $L_1$ nodes in the
% overlay.
% \item Any node can have children with level lower than theirs, except an
%  $L_{max}$ node which can only have $L_{max}$ level children and only in the
% case when its parent has reached its maximum degree.
% \item Each node is assigned a dot-formated address (e.g 1.3.4). Each octet
% ranges from
% $1$ to $d$, where $d$ is the maximum degree of the nodes. The assignment
% process
% is done by appending a unique octet to the address of each nodes parent, while
% root node is assigned address $1$.
% \item For any descendant node $Y$ of a node $X$, the measured distance among
%  each other, must always be less than the lower bound of the RTT interval
% denoted by $X$'s label.
%\end{enumerate}
%

The routing scheme is similar to the longest-prefix IP matching scheme. At each
forwarding hop, any message travels up the tree until the first common ancestor
of both source and destination nodes is reached and then starts descending to
arrive to its target. During tree traversals, special entries in the routing
tables, called \emph{routing cache}, are maintained in order to increase routing
efficiency and achieve finer load balance. Caching enables a node to forward a
message to a better longest-prefix match than that of its direct ancestor making
a larger, quicker and more cost effective step through the overlay and toward
the destination. To improve scalability, the number of children nodes and the
size of the routing cache are limited. In terms of overlay maintenance, LAPTOP
incorporates a simple \emph{heartbeat}-based technique where each parent node is
responsible for monitoring its children.

%
%The newcommer $N$ locates the root node and the later responds with a list of
%  $L_1$ nodes. $N$ then probes each of the $L_1$ nodes in search for the
% closest one. If the measured RTT to the closest $L_1$ falls into the first
% interval then the newcommer becomes a $L_1$ node as well. Otherwise node $N$
% sets the closest $L_1$ node as its potential parent node. This potential
% parent becomes the actual parent if it does not have any other children. If it
% has, $N$ gets a list of these $L_i$ nodes and by measuring the RTT to each of
% them tries to spot a new potential parent in order to repeat the above
% process.
%

% At join process the newcomming node is assigned its level label as well as its
% address by its parent node. Additionally, it initializes its routing table
%(with normal and caching entries) as it traverses the overlay in search for
%its parent node.  During a graceful departure, the referred node checks for
%children in the overlay. If it does not have any, it simply notifies its
%parent and leaves. If it has, it selects the child node with the lowest
%round-trip time to it in order to take its place so that the locality
%property is preserved. On the other hand, in the case of an arbitrary
%failure, after the children of the failed node detect their parent's absence
%(throught the aformentioned heartbeat mechanism), they start emmiting special
%messages to their grandparent node (the address of which is stored during
%join process}. In case of no response from the grandparent node, the children
%invoke the joining process to detect their new parent node.The parent of the
%failed node, aggregates the notifications from the above children nodes for a
%period of time, and then chooses the one with the lowest level label as the
%takeover node. Potential ties are broken by favouring the lower RTT. The
%parent of the failed node, finally informs accordingly all the children about
%the change in the hierarchy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% TODO: HERE STRUCTURED LANDMARK BINING APPROCH LIES
%
%Assuming a structured approach based on CAN\cite{ratnasamy_can_2001} and $m$
% landmark nodes. The coordinate space is then partitioned into $m!$ equally
%sized portions, each corresonding to a single ordering of the landmarks. To do
%this, the first dimension is divided into $m$ areas each of which is furtherly
%divided (second dimension) into $m - 1$ sections and so on. Having set this $m$
%dimensional space, at joining time, a node measures the delay to the set of
%landmarks in order to determine its associated bin and thus position itself in
%that portion of the coordinate space associated with its landmark ordering.
%Even though this scheme can
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Proximity Routing}
\label{section:pr}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Proximity routing does not require routing tables to be built using any
knowledge about network proximity. On the other hand it exploits such knowledge
in order to choose the best next hop during routing a message. This approach
balances between choosing the node that will further progress the routing
towards the destination and choosing the closest entry in the routing table, in
terms of network proximity.
% TODO: looks like discussion material
Thus, it is relatively less effective than
geographical layout when applied to CAN(-like) implementations. Moreover, the
technique has been incorporated into a version of Chord causing an increase on
the overhead of node joins and the size as well as maintenance cost of finger
tables.

% Moreover, the technique has been incorporated into a version of Chord
% (TODO: PChord | additionally check whether this is true) causing an increase
% on the overhead of node joins and the size as well as maintenance cost of
% finger tables.

%
%\cite{dabek_cfs_2001} proposes a server selection scheme for the Chord DHT, on
%the domain of proximity routing selection. In \emph{CFS}, each node predicts
%the entire lookup latency as a function of the total number of nodes and the
%average overlay next routing peer. The problem is that it is very difficult to
%have a clear picture on the total number of nodes and the average hop latency
%from the local. This leads to rough estimations that consequently decreases
%overall performance.
%
The methods that use proximity routing and their details are listed below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf Proximity in Kademlia}
% Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
% summarize the highlights about the {\em Kademlia} protocol. {\em Kademlia}
% \cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
% protocol for P2P networks, which uses an iteration based lookup algorithm. The
% protocol uses the standard $160$ bit ID system for nodes and locates the nodes
% in a prefix binary tree, where IDs are used as prefixes. The iterative lookup
% operations are done over this prefix binary tree, which converges to
%logarithmic
% lookup times. The ID's are assigned randomly, therefore in Kademlia, there is
%no
% proximity control, which results in inefficient use of the underlying network
% during lookup and retrieve operations.

\emph{Proximity in Kademlia} \cite{kaune_pkad_2008} introduces proximity
controls over the vanilla Kademlia protocol's random ID assignment, in order to
optimize the underlying network usage of the protocol. The work in this paper
can also be considered, as a more generic study over building proximity in the
context of iterative lookup algorithms. Focus is given on both the overlay and
the underlaying tiers in the sense that in the first improves the routing
performance according to a cost metric that is provided by the later and
quantifies the suitability of establishing a communication link between peers as
a cost function, based on the used proximity criteria (RTT, ISP locality, etc.).
%An underlay metric provides information about the underlay network. The paper
%  distingushes between three kinds of possibilities to quantify it:
%\begin{itemize}
% \item Using measurements gained by previous lookups
% \item Using measurements acquired by other peers or jointly calculated with
%  others
% \item Using a local database to look up information
%\end{itemize}
The paper proposes two approaches for overlay optimisation. One falling into
{\em proximity routing} and the other on {\em proximity neighbour selection}
(see Section~\ref{section:pns}) approaches overlay optimization techiques for
exploiting proximity. For the first, the goal is in choosing the best next hop
during routing a message. Due to the fact that the routing in Kademlia is
iterative, it is the initiator of the lookup that chooses each next hop from a
set of candidate nodes. Standard Kedemlia always chooses the closest node with
respect to the XOR metric but with proximity routing selection, Kademlia
chooses the one with the smallest underlay metric cost. Thus, as in all such
approaches, there is a trade-off between the overlay and underlay distances.

Toward the enhancement of traffic locality, the authors also use information
from the MaxMind GeoIP database\footnote{MaxMind Geolocation Technology.
http://www.maxmind.com} to retrieve proximity information of a given
peer based on its region, country, or ISP location, and use this information to
form clusters of nearby peers. An alternative method could also be the use of 
Vivaldi coordinates \cite{cox_vivaldi_2004}. However, authors report that
clustering approach performs better than Vivaldi protocol based on their
experiments. Authors also report that {\em Proximity routing} protocol
successfully worked with the Kademlia protocol and improved the locality of the
connections over peers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf CHOord considering Proximity on IPv6 (CHOP6)}
\emph{CHOP6} \cite{morimoto_chop6_2007} roughly estimates the proximity among
nodes by exploiting the IPv6 address format and RTT information if available.
The proximity estimation is achieved by introducing a 64-bit ID scheme in which
the least significant bit part is the IPv6 global routing prefix and thus
enabling a longest prefix match scheme. The protocol is designed based on the
observation that it is possible to estimate a node's geographical location by
simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
Chord, CHOP6 uses a finger table, whose entries hold more than
one candidate node.

%
%\paragraph{}
% There are three cases in which a node chooses the next hop according to
% information it posseses.
%\begin{itemize}
% \item When no information about RTT is available candidate next hops, the
%  sender just selects a node in the finger table entry which shares the longest
% prefix with the destination.
% \item There is another possibility that after some communication with other
%  nodes, the source node should know of the RTTs to some of the nodes in finger
% table entry. In this case, the source node chooses the one with the smallest
% RTT with propability $p$. If there is a node with no measured RTT then the
% sender can select such a node with probability $1 - p$
% \item In this last case the source node has already communicated with all
%  candidate node in the finger table entry. Thus, the node selects the node
% whose RTT is the smallest with probability $p$, the one with the second
% smallest with probability $q$ and so on, where $0 \leq \ldots < r < q < p < 1$
% and $p+q+r+\ldots \leq 1$
%\end{itemize}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{\bf T2MC}
T2MC \cite{shi_t2mc_2008} uses traceroute logs to detect different autonomous
systems and cluster close by peers, thus reducing redundant traffic that crosses
autonomous system boundaries. T2MC uses a customized k-mean classification
algorithm with $k=2$ to perform the classification, and exploits the stable
structure of the Internet routers to guide clustering. The peer chooses the
minimum and maximum Latency results from the traceroute to initialize the
centroinds of two sets. Then the peer calculates, for all hops allong its
tracerouted path, the absolute distance to these centroids of both sets and
assigns the routers to that centroid with which it has the smaller absolute
distance. After that, the mean and variance values are calculated and if the
variance is larger than a predefined threshold then the algorithm sets the two
latency mean values as new centroids for the two sets and loops again.
Ultimately peer will end up with two sets having the minimum intra-set variance.
The peer, then, chooses the router from the second set with the minimum hops
attribute and sets it as a threshold. The selected router and all others whose
hops attribute is larger than the threshold are classified as \emph{remote}
router cluster. The remaining are classified as \emph{near}. From the near
class, the peer chooses the one with maximum hops attribute as its edge router,
and registers it along with the all the near cluster into the DHT of the P@P
overlay. As new peers join the network, those that share the same edge router
or any of the members of the near router clustersm there would gather to form a
close peer cluster. As Edge routers can provide more valuable information than
other members of the near set, T2MC was designed to prioritize interaction of
peers and edge gateways.

%
% TODO: SOME DISCUSSION
%
%The use of Traceroute as a tool for implementing the distance measuring
%infrastructure raise concearns about its efficiency and scalability. Being,
%primarilly, a network diagnostic utility, it is concearned too heavy weighted
%and intrucive for use in a larger scale\cite{ratnasamy_binning_2002}.
%Additionally, disabling ICMP is a common administrative policy for edge sites
%to enforce security, while dumping BGP routing
%tables\cite{krishnamurthy_bgpclust_2000} is not directly available to the
%application layer.
% Even though traceroute provides detailed information about the network
% structure, its excessive use creates additional overhead that stresses further
% the overall network infrastructure. This forces network administrators to
% disable its support from the routers they manage, thus jeopardizing the
% effectiveness of the T2MC algorithm.
%
